Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-04-30 WSSSPE2_report/WSSSPE2_review.tex
Added notes on discussion group on reuse and sharing

In the first discussion session, this group broke up into two smaller groups working on different topics: Reproducibility, and Reuse and Sharing.

@@

>>>>

In the first discussion session, this group broke up into two smaller groups working on different topics: Reproducibility, and Reuse and Sharing. In the second discussion session, three groups worked on creating specific pieces of guidance.

@@

The second group discussed ways in which software could be more easily reused and shared. From the perspective of both user and developer, any solutions must be 1) easy; 2) cheap; 3) not take much time.

Reuse and sharing were considered to be distinct but linked. In many cases, pre-existing software does not exist, so new software is written but even then it is not shared afterwards. The principal barriers to reuse are the difficulty of finding out what software is available, understanding if it is usable, and then of installing and running software if it is located.

Discovery of relevant software is still fundamental issue: we need standard vocabularies and metadata, better tools, and approaches that are sustainable. Publications are an easy entry point for locating suitable software, but should the publishers lead the way, or is this the responsibility of "the community"? Some communities have had significant initiatives to improve software discovery, e.g. the NIH Software Discovery Index\footnote{NIH Software Discovery Index: http://softwarediscoveryindex.org/}. Likewise, there were examples of journals which had made software more discoverable: ACM Transactions on Software offer a reproducibility review; the Journal of Biostatistics has an opt-in to provide code and a certification mark if it can be run; the Journal of Open Research Software requires software to be deposited in suitable repositories and referenced.

An issue around the sustainability of software catalogues is that their usefulness often depends on the domain. For instance, in the bio-sciences there is more homogenous data, and more standard shred codes. In areas like ecology, codes are very specific to the problem, meaning that the level of re-use might be at a very general statistical level of abstraction, but then every research use is highly customized.

Finally, it was clear from discussion that there were many ways in which software could be made available in more reusable ways that just a tarball sitting on a personal website. Using code repositories like GitHub\footnote{GitHub: http://github.com/} gets you an archived, shared platform and improve the reusability of your software incrementally, for instance by adding a licence or by archiving (with a DOI) in Figshare\footnote{FigShare: http://figshare.com/} or Zenodo\footnote{Zenodo: http://zenodo.org/}. Docker\footnote{Docker: htp://www.docker.com/} might be a solution to the issue of dependencies, to allow binaries and libraries to be bundled in a more lightweight fashion than a virtual machine image.

The key enabler for reuse and sharing was to get domain scientists more effectively connected with programmers/analysts. Both have skills and experience which is necessary to make the right decisions for improving the reusability and discoverability of software, and to apply community pressure to change practice.

\subsubsection{Categorisation of software journals}

\subsubsection{What should journals provide reviewers of software papers}

\subsubsection{Reproducibility Meta-track toolkit}
